\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}

% ---- result macros: fill from bench/results/*.json once n=50 runs land ----
% BUC (belief update) — Stale Leakage Rate / Update Accuracy
\newcommand{\bucSLRfull}{\textbf{34.0}}   % nemos-v2-semantic
\newcommand{\bucSLRvone}{50.0}            % nemos-v1-lexical (shipped)
\newcommand{\bucSLRnone}{80.0}            % no-invalidation (floor)
\newcommand{\bucUAfull}{76.0}
\newcommand{\bucUAvone}{84.0}
\newcommand{\bucUAnone}{92.0}
% mem0 external reference (BUC) — mem0 v2.0.8, 50 items, shared judge
\newcommand{\bucSLRmemo}{82.0}
\newcommand{\bucUAmemo}{96.0}
% ASP (anti-self-pollution) — Pollution Rate / User-Fact Recall  [n=50 items, 250 probes]
\newcommand{\aspPRiso}{\textbf{1.6}}      % isolated (Nemos)
\newcommand{\aspPRshared}{71.2}           % shared store (baseline)
\newcommand{\aspUFRiso}{92.4}
\newcommand{\aspUFRshared}{92.0}
% FOR (forgetting) — Trivia Leakage / Important-Fact Retention  [n=50 items, 149 probes]
\newcommand{\forTLdecay}{\textbf{17.4}}
\newcommand{\forTLnone}{96.0}
\newcommand{\forIFRdecay}{75.2}
\newcommand{\forIFRnone}{73.8}
\newcommand{\nItems}{50}                  % items per task
% LongMemEval knowledge-update slice (external cross-anchor), QA accuracy
\newcommand{\lmeAccFull}{\textbf{43.3}}   % nemos-v2-semantic
\newcommand{\lmeAccNone}{33.3}            % no-invalidation
\newcommand{\lmeN}{30}

\title{MnemoBench: Evaluating Belief Update, Self-Pollution Resistance,\\
and Forgetting in Long-Lived Memory Systems}
\author{Shen Wei\thanks{Working draft. All reported numbers are from the frozen
$n{=}50$ benchmark runs in the released harness.}\\
\texttt{s181@163.com}\\
Project: \url{https://github.com/mmlong818/nemos}}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
Persistent memory layers for LLM agents are typically evaluated on \emph{recall}:
can the system retrieve what was said earlier? We argue that in long-lived use the
dominant failure modes are not recall but \emph{maintenance}: (1) failing to revise a
belief when a fact changes, (2) letting the agent's own generated or imagined content
pollute the user's fact base, and (3) never forgetting, so that stale trivia degrades
precision over time.
We introduce \textbf{MnemoBench}, a reproducible benchmark with
three task families targeting exactly these behaviours, with ground truth fixed by a
generator rather than judged post~hoc. We evaluate Nemos, an embeddable memory kernel
with bitemporal contradiction invalidation, namespace isolation between user facts and
agent self-narrative, and FSRS-based decay. On belief update, contradiction invalidation
cuts stale-answer leakage from \bucSLRnone{}\% to \bucSLRfull{}\%, at a moderate recall
cost (update accuracy \bucUAnone{}~$\rightarrow$~\bucUAfull{}\%) that exposes a
precision/recall knob; we further show that the shipped lexical contradiction detector
misses attribute-replacement updates and that a semantic detector recovers them
(leakage \bucSLRvone{}~$\rightarrow$~\bucSLRfull{}\%). Namespace isolation reduces
self-pollution from \aspPRshared{}\% to \aspPRiso{}\% with no recall loss. Decay suppresses
stale-trivia leakage from \forTLnone{}\% to \forTLdecay{}\% while retaining important facts.
We release MnemoBench and all harness code.
\end{abstract}

\section{Introduction}
\label{sec:intro}
% The gap: memory benchmarks measure recall; real long-term failure is maintenance.
Memory systems for LLM agents (e.g.\ MemGPT/Letta~\cite{memgpt}, mem0~\cite{mem0},
Zep/Graphiti~\cite{zep,graphiti}) are commonly assessed on multi-session question answering
such as LoCoMo~\cite{locomo} and LongMemEval~\cite{longmemeval}, which primarily stress
\emph{retrieval}. Yet the failures that erode trust in a long-lived assistant are different:
\begin{itemize}
  \item \textbf{Belief staleness.} A fact changes (job, city, diet, relationship) and the
  system keeps surfacing the old value as if current --- a failure mode recent work on
  knowledge updating in memory systems highlights as both common and hard~\cite{fama}.
  \item \textbf{Self-pollution.} In companion/agent settings the model emits first-person
  or speculative content; if mis-attributed it enters the user profile and is later returned
  as the user's own authoritative fact.
  \item \textbf{No forgetting.} Append-only stores accumulate one-off trivia that crowds out
  salient facts and lowers retrieval precision.
\end{itemize}
We make two contributions. (i) \textbf{MnemoBench}, a benchmark isolating these three
behaviours with generator-fixed ground truth (\S\ref{sec:bench}). (ii) An evaluation of the
\textbf{Nemos} memory kernel (\S\ref{sec:system}) showing each mechanism is responsible for a
measurable, ablatable improvement (\S\ref{sec:results}), including a before/after on a real
defect we found and fixed in the contradiction detector.

\section{Related Work}
\label{sec:related}
\paragraph{Memory architectures.} MemGPT/Letta~\cite{memgpt}, mem0~\cite{mem0},
Zep/Graphiti's temporal knowledge graph~\cite{zep,graphiti}. Contrast: layered separation,
explicit invalidation semantics, source-authority gating.
\paragraph{Temporal/ bitemporal modelling.} Graphiti edge validity intervals~\cite{graphiti}
vs.\ Nemos event-sourced invalidation with materialised flat fields.
\paragraph{Forgetting.} FSRS / spaced-repetition retrievability~\cite{ssp,fsrs} applied to
machine memory rather than human review scheduling.
\paragraph{Benchmarks.} LoCoMo~\cite{locomo} and LongMemEval~\cite{longmemeval} are
recall-centric. Closest to our thesis is the concurrent work of Uddin et al.~\cite{fama},
whose \emph{Memora} benchmark and Forgetting-Aware Memory Accuracy (FAMA) metric explicitly
penalise reliance on obsolete or invalidated memory --- corroborating the
recall-to-maintenance shift. MnemoBench is complementary: rather than a single aggregate
score, it isolates each maintenance mechanism (invalidation, isolation, decay) through
single-mechanism ablations, and additionally surfaces a before/after on a real detector fix.

\section{The Nemos Memory Kernel}
\label{sec:system}
Nemos is an embeddable memory kernel. We describe only the mechanisms under test.
\subsection{Layered memory and source authority}
Five layers (episodic, semantic, personal\_semantic, procedural, archival). Every record
carries \texttt{source.authoritative} (user-stated vs.\ model-derived); derived content is
barred from authoritative personal facts (invariant I4).
\subsection{Bitemporal contradiction invalidation}
Records carry world-axis (\texttt{valid\_at}/\texttt{invalid\_at}) and system-axis
(\texttt{created\_at}/\texttt{expired\_at}) timestamps and a
\texttt{belief\_state}~$\in$~\{active, invalidated, superseded, corrected\}. During offline
reflection, a new fact that contradicts an existing \texttt{personal\_semantic} anchor marks
the old one \texttt{invalidated}; default retrieval returns only \texttt{active} records.
\paragraph{A real defect: lexical vs.\ semantic detection.}
The shipped detector prefilters contradiction candidates by character-bigram Jaccard
similarity. This reliably catches strong negations (``left Google'') but \emph{misses
attribute replacement} where the new value is lexically dissimilar (vegetarian
$\rightarrow$ pescatarian). We replace the prefilter with embedding-cosine candidate
retrieval and add a same-attribute, mutually-exclusive reconciliation pass over existing
personal facts. \S\ref{sec:results} quantifies the before/after.
\subsection{Namespace isolation (anti-self-pollution)}
User facts live under \texttt{forUser(uid)}; agent self-narrative under
\texttt{forUser(persona:pid)} --- a physically separate keyspace. Retrieval of user facts
cannot surface persona content.
\subsection{FSRS decay}
Each record has stability $S$, difficulty $D$, retrievability
$R=\exp(-\Delta t / S)$. Access reinforces $S$; records with $R$ below a cold threshold,
zero recent access, after a dormancy window are marked cold and hidden from default search
(archival is never cooled).

\section{MnemoBench}
\label{sec:bench}
\subsection{Design principles}
Ground truth is \emph{fixed by the generator}: each item's fact script (which attribute,
which values, in what order; which persona statements are traps) is decided first, then
rendered into natural-language sessions. Scoring isolates \emph{memory quality} from
\emph{generation quality}: an LLM judge inspects only the \emph{set of retrieved records}
and decides whether the expected fact is present and whether a forbidden (stale/leaked)
fact is presented as current. The judge never writes the final answer.
\subsection{Task families}
\paragraph{BUC --- Belief Update \& Contradiction.} A single-valued user attribute changes
1--3 times across sessions; the probe asks for the current value.
\textit{Update Accuracy} (expected present) and \textit{Stale Leakage Rate} (a prior value
presented as current; lower better).
\paragraph{ASP --- Anti-Self-Pollution.} User statements interleaved with persona
self-statements, some crafted as plausible traps. Probes query user facts.
\textit{Pollution Rate} (persona fiction returned as authoritative user fact; lower better)
and \textit{User-Fact Recall}.
\paragraph{FOR --- Forgetting \& Salience.} Durable important facts mixed with one-off
trivia at older timestamps; salient facts are revisited during the retention interval.
\textit{Trivia Leakage} (cold trivia still surfacing; lower better) and
\textit{Important-Fact Retention} (guardrail; must stay high).
\subsection{Item schema and scoring}
Each item is a JSON record with an ordered list of \texttt{sessions}
(\texttt{\{speaker, text, contentDate?\}}) and one or more \texttt{probes}
(\texttt{\{query, expected[], forbidden[]\}}). For BUC the labels are derived
deterministically from the fact-script (the last value is the only \texttt{expected}
current value; all earlier values are \texttt{forbidden}), so the labels cannot drift from
the rendered narrative. The judge receives the query, the \texttt{expected} and
\texttt{forbidden} lists, and the retrieved record set, and returns two booleans ---
\texttt{contains\_expected} and \texttt{contains\_forbidden} --- where a record that
explicitly marks a forbidden value as past does \emph{not} count as a leak.

\section{Experimental Setup}
\label{sec:setup}
% Models, ablation arms, n, judge model, reproducibility.
All systems use OpenAI \texttt{gpt-4o} for extraction/reflection and
\texttt{text-embedding-3-small} for embeddings; the judge is \texttt{gpt-4o}. We report
$n=\nItems{}$ items per task. Each task is evaluated by ablating exactly the mechanism under
test, so any difference is attributable to that mechanism:
\begin{itemize}
  \item BUC: \texttt{semantic} (full) vs.\ \texttt{lexical} (shipped) vs.\ invalidation-off.
  \item ASP: namespace-isolated vs.\ shared store.
  \item FOR: decay-on vs.\ decay-off. To isolate decay we disable reflection in both arms and
  use cold threshold $0.3$, dormancy $0$, with a 120-day retention interval; important facts
  are accessed during the interval (modelling that salient facts are revisited).
\end{itemize}
As one external reference point we run mem0 (v2.0.8) on the BUC task with the same models and
the identical judge. As a standard-benchmark cross-anchor we additionally evaluate a
\lmeN{}-question slice of the LongMemEval~\cite{longmemeval} \emph{knowledge-update} subset
(oracle setting), following its generate-then-judge QA-accuracy methodology, ablating
invalidation (semantic vs.\ off).

\section{Results}
\label{sec:results}
\begin{table}[t]\centering
\caption{Belief update (BUC). Lower SLR is better; UA is the recall guardrail.}
\begin{tabular}{lcc}
\toprule
System & Update Acc.\ (\%) & Stale Leakage (\%) \\
\midrule
mem0 (external)\footnotemark{}      & \bucUAmemo{} & \bucSLRmemo{} \\
No invalidation (floor)        & \bucUAnone{} & \bucSLRnone{} \\
Lexical detector (shipped)     & \bucUAvone{} & \bucSLRvone{} \\
Semantic detector (this work)  & \bucUAfull{} & \bucSLRfull{} \\
\bottomrule
\end{tabular}
\end{table}
\footnotetext{mem0 v2.0.8, same gpt-4o + text-embedding-3-small, scored by the identical judge.}

\begin{table}[t]\centering
\caption{Anti-self-pollution (ASP) and forgetting (FOR).}
\begin{tabular}{lcc}
\toprule
\multicolumn{3}{l}{\emph{ASP} --- Pollution Rate / User-Fact Recall (\%)}\\
\midrule
Shared store (baseline)  & \aspPRshared{} & \aspUFRshared{} \\
Namespace isolation      & \aspPRiso{}    & \aspUFRiso{} \\
\midrule
\multicolumn{3}{l}{\emph{FOR} --- Trivia Leakage / Important Retention (\%)}\\
\midrule
Decay off                & \forTLnone{}   & \forIFRnone{} \\
Decay on                 & \forTLdecay{}  & \forIFRdecay{} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Belief update (BUC).} Stale leakage falls monotonically across the ablation
ladder: \bucSLRnone{}\% with no invalidation, \bucSLRvone{}\% with the shipped lexical
detector, and \bucSLRfull{}\% with the semantic detector --- the contradiction mechanism more
than halves stale answers, and the lexical$\rightarrow$semantic change alone removes a further
third of the leakage by catching attribute-replacement updates the lexical prefilter misses.
This gain is not free: update accuracy drops from \bucUAnone{}\% (floor) to \bucUAfull{}\%
(semantic), i.e.\ aggressive invalidation occasionally retires a still-current fact. BUC thus
exposes a precision/recall knob absent from the other two tasks; we regard the operating point
as tunable (candidate-similarity threshold, reconciliation conservativeness) rather than fixed.
The external system, mem0, retains contradicted facts as active alongside their replacements
(it appends without invalidating the old), placing its leakage near the no-invalidation floor.

\paragraph{External cross-anchor (LongMemEval).} The same effect appears on a standard
benchmark: on a \lmeN{}-question slice of the LongMemEval knowledge-update subset (QA accuracy,
generate-then-judge), enabling semantic invalidation raises accuracy from \lmeAccNone{}\% to
\lmeAccFull{}\% --- a $+10$-point gain attributable solely to the invalidation ablation. The
absolute level is modest (Nemos is used as a from-scratch memory over the oracle evidence
sessions, without LongMemEval-specific tuning), so we read the \emph{delta}, not the level, as
the result: belief maintenance helps even on recall-style external data.

\paragraph{Anti-self-pollution (ASP).} This is the largest and cleanest effect. Across
$250$ user-fact probes, namespace isolation reduces the pollution rate from \aspPRshared{}\%
(a single shared store, as in append-only designs) to \aspPRiso{}\% --- a $\sim$44$\times$
reduction --- while user-fact recall is statistically unchanged (\aspUFRshared{}\% vs.\
\aspUFRiso{}\%). Physically separating the agent's self-narrative keyspace from the user's
facts essentially eliminates the failure mode at no cost to recall.

\paragraph{Forgetting (FOR).} Over $149$ salience probes, decay cuts trivia leakage from
\forTLnone{}\% to \forTLdecay{}\% while important-fact retention is unchanged
(\forIFRnone{}\% vs.\ \forIFRdecay{}\%): unreferenced trivia is cooled out of default
retrieval without dropping the revisited important facts.

\section{Discussion and Limitations}
\label{sec:limits}
\paragraph{LLM-judge dependence.} Both rendering and scoring use an LLM; we mitigate with a
set-membership judge (no answer generation), low temperature, and a human-audited sample.
\paragraph{Self-generated data.} Items are synthetic; we report only relative differences
under identical data and add a standard-benchmark slice (LongMemEval) as a cross anchor.
\paragraph{FOR isolation choices.} Disabling reflection and using aggressive decay parameters
isolate decay cleanly but do not reflect the full-system default; we report them explicitly.
\paragraph{Residual leakage.} The semantic detector does not reach zero stale leakage on
generated hard cases (lexically and semantically close value pairs); we analyse failures in
\S\ref{sec:results}.

\section{Conclusion}
Long-lived memory should be judged on maintenance, not just recall. MnemoBench operationalises
this, and each Nemos maintenance mechanism yields a measurable, ablatable gain.

\bibliographystyle{plain}
\bibliography{refs}
\end{document}
